---
title: "Beer Production data"
author: "Alexandra Akaakar"
output:
  html_document:
    css: "style.css"
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE , message = FALSE, warning = FALSE)
```

## About This Project

This is an analysis of Beer Production data gotten from the Alcohol and Tobacco Tax and Trade Bureau.

The Datasets include information on:
State-level beer production by year 

Number of brewers by production size by year 

Monthly beer stats aggregated across the US 

Some considerations:

A barrel of beer for this data is 31 gallons.

Most data is in barrels removed/taxed or produced.

Removals = "Total barrels removed subject to tax by the breweries comprising the named strata of data". Essentially how much was produced and removed for consumption.


```{r}
#load libraries
library(tidyverse)
library(lubridate)
library(ggplot2)
library(dplyr)
library(plotly)
library(hrbrthemes)
library(magrittr)
library(remotes)
```


```{r}
#Get data
brewing_materials <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-31/brewing_materials.csv')%>%
mutate(date = ymd(paste(year, month, 1)))
#%>% filter(year == 2017)

beer_taxed <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-31/beer_taxed.csv') 
#%>% filter(year == 2017)


brewer_size <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-31/brewer_size.csv') 
#%>% filter(year == 2017)


beer_states <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-31/beer_states.csv') 
#%>% filter(year == 2017)

```
 
 delete these later
```{r,include=FALSE}
#checking dimensions
dim(brewing_materials)
dim(beer_taxed)
dim(brewer_size)
dim(beer_states)
```

```{r,include=FALSE}
#summary and column names

colnames(brewing_materials)
summary(brewing_materials)


colnames(beer_taxed)
summary(beer_taxed)


colnames(brewer_size)
summary(brewer_size)

colnames(beer_states)
summary(beer_states)

```

## Brewing Materials dataset Analysis

What type of materials are used in US Beer products?

The Brewing Materials dataset contains information on the materials that went into beer produced in the united states 
the column names in the dataset are:
```{r}
#summary and column names
colnames(brewing_materials)


#checking dimensions
dim(brewing_materials)

```


material_type shows the number of pounds of Hops vs Grains that went into Beer production.
Type expands on material_type.

month_current shows how many barrels of Beer were produced for this year/month. Month_prior_year shows how many barrels were produced at the same period last year.

ytd_current shows the cumulative production in barrels for the current year.  ytd_prior_year shows the cumulative production in barrels for same period last year.


```{r}


grain_nongrain_input <- as.tibble( 
  brewing_materials %>% group_by(year,month,material_type,type)%>% summarise(sum(month_current))%>%
    rename(Monthlyamt = `sum(month_current)`)
  ) 
```


```{r}
## Boxplot of materils
grain_nongrain_input %>% ggplot(mapping = aes(Monthlyamt,month),fill = type )+
  geom_boxplot(mapping = NULL,
  data = NULL,
  stat = "boxplot",
  position = "dodge",
  outlier.colour = "#3366FF",
  outlier.color = NULL,
  outlier.fill = NULL,
  outlier.shape = 19,
  outlier.size = 1.5,
  outlier.stroke = 0.5,
  outlier.alpha = NULL,
  notch = TRUE,
  notchwidth = 0.5,
  varwidth = TRUE,
  na.rm = FALSE,
  orientation = NA,
  show.legend = NA,
  inherit.aes = TRUE) +
  facet_wrap(~year)+
  coord_flip()
  
```
```{r}
materialcountyear <- as.tibble( 
  brewing_materials %>% group_by(year,material_type)%>% summarise(sum(year_current))
)
```

```{r}
ggplot(brewing_materials, aes(type,fill = material_type)) +
  fct_reorder(material_type, tvhours)+
  geom_bar()+
  coord_flip()+
  facet_wrap(~year)

#need to relevel factors and add

```


```{r}
colSums(sapply(brewing_materials, is.na))

brewing_materials <- as.data.frame(unclass(brewing_materials), stringsAsFactors = TRUE)

str(brewing_materials)

```

```{r}
class(brewing_materials)
summary(brewing_materials)

#The %<>% operator pipes and reassigns, which is very useful for keeping data cleaning and transformation simple. it requires the maagrittr function
#brewing_materials[1:2] %<>% lapply(function(brewing_materials) as.character(as.factor(brewing_materials)))

```


The Brewing Material dataset covers Material used in Beer Production from 2008 to 2017.
It has 144 rows and 10 columns. YTD_current and YTD_prior_year have 144 missing observations.



```{r}
month_levels <- c(
  "Jan", "Feb", "Mar", "Apr", "May", "Jun", 
  "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"
)


brewing_materials %>% mutate() <- factor(brewing_materials$month, levels = month_levels)
```





dataset numerical distribution:
```{r}

materialnumbers <- brewing_materials %>% select(where(is.numeric))

materialtype <- brewing_materials %>% select(c(year, month ,material_type,type))

str(materialnumbers)
str(materialtype)
```


```{r}
materialnumbers %>% ggplot(x = month) +
  geom_col(mapping = aes( y = (month_current), 
           position = 'stack') + 
  facet_wrap( ~ year)

ggplot(data = brewing_materials) + 
  geom_col(mapping = aes(x = month, y = (month_prior_year/1000000), fill = month)) + 
  facet_wrap(~year)

ggplot(data = brewing_materials) + 
  geom_col(mapping = aes(x = month,y = (ytd_current/1000000)), fill = month) + 
  facet_wrap(~year)

ggplot(data = brewing_materials) + 
  geom_col(mapping = aes(x = (ytd_prior_year/1000000), y = month)) + 
  facet_wrap(~year)

```

The distribution of the production numbers in the Beer_materials dataset are all right-skewed. The mean median and mode differ.

```{r}

BrewMatNum <- brewing_materials %>% select(month_current,month_prior_year,ytd_current,ytd_prior_year,year) %>% 
filter(!is.na(ytd_current), !is.na(ytd_prior_year)) %>% group_by(year)

do.call(cbind, lapply(BrewMatNum, summary))
```

```{r}
<!-- ```{r,include=FALSE} -->
<!-- p4 <- hist(brewing_materials, col="#F78181", xlab="Rent", main="Distribution of Rent") -->
<!-- abline(v = mean(subset.rent$rent_mean), col = "blue", lwd = 2, lty="dashed") -->
<!-- abline(v = median(subset.rent$rent_mean), col = "green", lwd = 2, lty="dashed") -->
<!-- legend(x = c(4000, 3200), y = c(8000, 5500), legend=c("Mean", "Median"), col=c("blue","green"), cex=0.7,  -->
<!--       lty="dashed", lwd=1, y.intersp = 3.8, x.intersp=3.5, xjust=-1.8) -->
```

material type analysis

```{r}

#materialtype <- brewing_materials %>% select(c(year, month ,material_type,type))
head(materialtype)

materialtype %>% ggplot(aes(month, material_type, fill = type)) +
  geom_col()+
  facet_wrap(~year)
  coord_flip()

```


```{r}

```

