---
title: "maps and spatial data"
author: "Alexandra Akaakar"
date: "10/25/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.width = 7, fig.height = 7)
loadLibs = function(libs="RgoogleMaps"){
  for (l in libs) 
    suppressMessages(suppressWarnings(library(l,character.only =TRUE, quietly=TRUE)))
}
loadLibs("RgoogleMaps")
loadLibs("loa")
loadLibs("latticeExtra")
loadLibs("ggmap")
library(ggplot2)
library(maps)
library(mapdata)


# bingAPIkey = scan("C:/Users/loecherm/Dropbox/stuff/bingAPIkey.txt",what="")
# bingAPIkey = scan("C:/DatenLoecher/Dropbox/stuff/bingAPIkey.txt",what="")
# googleAPIkey = scan("C:/DatenLoecher/Dropbox/stuff/googleAPIkey.txt",what="")
# dataDir="H:/DropboxHWR/InternationalVisits/Aalborg2016/Teaching/MapsInR/ForStudents/data/"
# OFFLINE=TRUE # use previously downloaded maps/data
# set.seed(123)
# if (OFFLINE) load("data/allTiles.rda")
```

Geographic Information Systems (GIS) is a powerful field that combines programming, geospatial data analysis, and visualization. 
R offers several packages to work with GIS data effectively. Here are some fundamental skills you can learn in R to leverage your programming:

1. Data Handling & Manipulation: Familiarize yourself with reading and writing spatial data formats like shapefiles (using the 'sf' package) or GeoJSON (with the 'jsonlite' package). Learn how to manipulate, clean, and aggregate geospatial data using packages such as dplyr and tidyr.

2. Spatial Data Visualization: Master essential visualization techniques for mapping with R using libraries like ggplot2 or leaflet. This includes creating choropleth maps (mapping quantitative variables onto geographic regions), scatterplots, heatmaps, and more.

3. Geostatistical Analysis & Modeling: Learn how to perform spatial regression analysis, interpolation techniques (kriging with the 'geoR' package), or cluster analysis using libraries like 'spatstat'. This will enable you to extract insights from geospatial data effectively.

4. Spatial Data Processing & Analysis: Understand how to work with raster and vector data, perform overlay operations (e.g., buffer zones, intersection, union), and calculate spatial statistics using packages like 'raster' or 'rgdal'.

5. Geocoding & Reverse Geocoding: Learn how to convert addresses into geographic coordinates (geocoding) and vice versa (reverse geocoding) with the help of libraries such as 'ggmap', 'osmdata', or 'geocoder'.

6. Network Analysis: If you need to work with network data, learn how to analyze road networks, transportation systems, or social networks using packages like 'igraph' or 'sna'.

7. Time Series & Dynamic Maps: Learn how to create dynamic maps that update over time (e.g., animated choropleth maps) using libraries such as 'mapview', 'animar', and 'sf'.

8. Integration with GIS Software: Familiarize yourself with tools like QGIS, ArcGIS, or Google Earth Engine to integrate your R scripts into a broader GIS workflow.

9. Spatial Data Management & Organization: Organize your geospatial data using databases (e.g., PostgreSQL) and spatial extensions like PostGIS. Learn how to manage large datasets efficiently with the help of libraries such as 'fossil' or 'dbRipple'.

10. Collaboration & Open Source: Contribute to open-source GIS projects in R, attend meetups, or join online communities (e.g., rOpenSci) to expand your knowledge and skills while helping the broader community.

By mastering these fundamental GIS skills in R, you'll be able to leverage your programming expertise for geospatial data analysis and visualization effectively.

```{r}
library(RCurl)
library(RJSONIO)
library(plyr)
```
first we Build a URL to access the API:

```{r}
url <- function(address, return.call = "json", sensor = "false") {
  root <- "http://maps.google.com/maps/api/geocode/"
  u <- paste(root, return.call, "?address=", address, "&sensor=", sensor, sep = "")
  return(URLencode(u))
}

```


to parse results

```{r}
geoCode <- function(address,verbose=FALSE) {
  if(verbose) cat(address,"\n")
  u <- url(address)
  doc <- getURL(u)
  x <- fromJSON(doc,simplify = FALSE)
  if(x$status=="OK") {
    lat <- x$results[[1]]$geometry$location$lat
    lng <- x$results[[1]]$geometry$location$lng
    location_type  <- x$results[[1]]$geometry$location_type
    formatted_address  <- x$results[[1]]$formatted_address
    return(c(lat, lng, location_type, formatted_address))
    Sys.sleep(0.5)
  } else {
    return(c(NA,NA,NA, NA))
  }
}
```


test your setup with an address

```{r}
address <- geoCode("The White House, Washington, DC")
address
```



```{r, echo=TRUE, eval=!OFFLINE, include=FALSE}
#par(pty="s")
#map1 <- GetMap(markers = paste0(
#"&markers=color:blue|label:S|40.702147,-74.015794",
#"&markers=color:green|label:G|40.711614,-74.012318",
#"&markers=color:red|color:red|label:C|40.718217,-73.998284"),
#center=NULL,destfile = "staticMaps/MyTile1.png",size=c(640,640),
#maptype = "terrain", API_console_key=googleAPIkey)
#PlotOnStaticMap(map1)
```
```{r}
w2hr <- map_data("world2Hires")
w2hr
usa <- map_data("usa")

dim(usa)
```

 database structure
The structure of those data frames
These are pretty straightforward:

long is longitude. Things to the west of the prime meridian are negative.
lat is latitude.
order. This just shows in which order ggplot should “connect the dots”
region and subregion tell what region or subregion a set of points surrounds.
group. This is very important! ggplot2’s functions can take a group argument which controls (amongst other things) whether adjacent points should be connected by lines. If they are in the same group, then they get connected, but if they are in different groups then they don’t.
Essentially, having to points in different groups means that ggplot “lifts the pen” when going between them. 
 
to create a map with with no line color and a black fill use geom_polygon()

```{r}

usa <- map_data("usa") # we already did this, but we can do it again
ggplot() + geom_polygon(data = usa, aes(x=long, y = lat, group = group)) + 
  coord_fixed(1.3)


```

```{r}
ggplot() + 
  geom_polygon(data = usa, aes(x=long, y = lat, group = group), fill = NA, color = "red") + 
  coord_fixed(1.3)
```

```{r}
map_data("state")
```

```{r}
mapImageData1 <- get_map(location = c(lon = -0.016179, lat = 51.538525),
    color = "color",
    source = "google",
    maptype = "satellite",
    zoom = 17)
 
ggmap(mapImageData1,
    extent = "device",
    ylab = "Latitude",
    xlab = "Longitude")
```

```{r}
?register_google
```

